---
output: github_document
header-includes: |
    \-\-\-  
    title: Usage with R Markdown documents  
    \-\-\-
---    

```{r setup, include=FALSE}
library(sketch)
knitr::opts_chunk$set(eval = FALSE)
knitr::knit_engines$set(sketch = eng_sketch)
```

## Embed a sketch R file in an R Markdown document

Suppose the sketch R file is `sketch.R`, then it can be embedded in an RMD file using `sketch::insert_sketch` in a code chunk. This is particularly useful when you have a large application.

```{r, eval=F}
sketch::insert_sketch(file = "sketch.R", style = "width:530px; height:220px; border:none;")
```

```{r, eval=T, echo=F}
sketch::insert_sketch(file = "sketch.R", style = NULL,
                      asset_tags = default_tags(local = FALSE),
                      class = "width-530 height-220 border-none")
```



## Use the sketch knitr engine

**sketch** now has a knitr engine! To use it,

- include `knitr::knit_engines$set(sketch = sketch::eng_sketch)` in the setup code chunk, and then
- use ```` ```{sketch}``` ```` in place of ```` ```{r}``` ````.

Concretely, the code chunks

````
`r ''````{r setup, include=FALSE}
knitr::opts_chunk$set(eval = FALSE)
knitr::knit_engines$set(sketch = sketch::eng_sketch)
```

`r ''````{sketch, debug = T, style = "width: 100%; height: 90px; border: none;"}
print("'sketch' has its own knitr engine from version 1.0.5!")
```
````

will produce

```{sketch, debug=magrittr::set_attr(TRUE, "local", FALSE), style=NULL, class="border-none width-full height-90 bg-white rounded-lg", eval=T, echo=F, asset_tags=default_tags(local = FALSE)}
print("'sketch' has its own knitr engine from version 1.0.5!")
```

The engine supports the following options:

- `rules`, `deparsers` to pass on to `source_r`,
- `debug`, `asset_tags` to pass on to `compile_r`, and
- `style` for styling the iframe, e.g., `style = "width: 100%; border: none;"` is a common option to use.


<link href="../css/docs-styles.css" rel="stylesheet"></link>
