---
output: github_document
header-includes: |
    \-\-\-  
    title: Creative coding with p5.js  
    slug: /  
    keywords:  
        - creative coding  
        - p5.js  
    \-\-\-
---    

```{r setup, include=FALSE}
library(sketch)
knitr::opts_chunk$set(eval = FALSE)
knitr::knit_engines$set(sketch = eng_sketch)
```

Created: 2021-04-03.
Modified: 2021-04-04.

## Introduction

In this article, we will do some basic creative coding together! We will create the following visualisation:

```{r, echo = F, eval = TRUE}
sketch::insert_sketch("creative_coding.R", 
                      asset_tags = default_tags(local = FALSE),
                      style = NULL,
                      class = "border-none width-420 height-420")
```

The first step to create such visualisation is to describe the visualisation using primitive operations. In this case, the visualisation is simply a 2-dimensional array of circles with randomised size and color. This gives us a 2-step plan for the implementation:

1. Draw a 2-dimensional grid of circles
2. Randomise the color and size of the circles


## Step 1 - Draw a 2-dimensional grid of circles

We begin by loading the **p5.js** library, setting up the coordinates where the circles will locate, then drawing the circles on the screen. 

Recall from the ["Getting Started"](/docs/) example that 

- [`setup`](https://p5js.org/reference/#/p5/setup) is a function **p5** monitors and calls once when your application launches,
- [`createCanvas`](https://p5js.org/reference/#/p5/createCanvas) is a function to create a canvas on the screen, and
- [`circle`](https://p5js.org/reference/#/p5/circle) is a function to draw a circle on the screen.

```{sketch, eval = TRUE, asset_tags = default_tags(local = FALSE), class = "border-none width-420 height-420", style = NULL}
#! config(deparsers = default_2_deparsers())
#! load_library("p5")

# Setup
WIDTH <- 400
HEIGHT <- 400
DIAMETER <- 10
GRID_SIZE <- 40

x_grid <- seq(0, WIDTH, step = DIAMETER)
y_grid <- seq(0, HEIGHT, step = DIAMETER)

setup <- function() {
    createCanvas(WIDTH, HEIGHT)
    for (x in x_grid) {
        for (y in y_grid) {
            circle(x, y, DIAMETER)
        }
    }
    NULL
}
```

Note that in the above we returned `NULL` at the end of `setup` because JavaScript does not allow returning assignments and control-flow statements (e.g. for-loop and if-statement). If you missed that, **sketch** would return `NULL` by default, and in this case the application would run fine; however, you will get a warning from **sketch**. In general, it is good practice to be explicit about what to return as there are situations where returning `NULL` may not be the intention.


## Step 2 - Randomise the color and size of the circles

Next, we randomise the color and size of the circles by

- using `fill` to change the fill color of the circles,
- using `stroke` to change the border color of the circles,
- modifying the diameter to change the size of the circles.

```{sketch, eval = TRUE, asset_tags = default_tags(local = FALSE), class="border-none width-420 height-420", style = NULL}
#! config(deparsers = default_2_deparsers())
#! load_library("p5")

# Setup
WIDTH <- 400
HEIGHT <- 400
DIAMETER <- 10
GRID_SIZE <- 40

x_grid <- seq(0, WIDTH, step = DIAMETER)
y_grid <- seq(0, HEIGHT, step = DIAMETER)

setup <- function() {
    createCanvas(WIDTH, HEIGHT)
    for (x in x_grid) {
        for (y in y_grid) {
            # Uniform distribution for RGB color
            rgb_color = runif(3, 0, 255)
            fill(rgb_color)    # fill color
            stroke(rgb_color)  # border color
            
            # Square of normal distribution for the diameter
            d <- DIAMETER * rnorm(1, 0, 0.4)^2
            circle(x, y, d)
        }
    }
    NULL
}
```

## Extra!

Let's add an extra feature that when we click on the canvas, it will regenerate another drawing! To do that, we wrap the drawing code into a function `draw_grid`, then add it into the [`mouseClicked`](https://p5js.org/reference/#/p5/mouseClicked) function.

```{sketch, eval = TRUE, asset_tags = default_tags(local = FALSE), class="border-none width-420 height-420", style = NULL}
#! config(deparsers = default_2_deparsers())
#! load_library("p5")

# Setup
WIDTH <- 400
HEIGHT <- 400
DIAMETER <- 10
GRID_SIZE <- 40

x_grid <- seq(0, WIDTH, step = DIAMETER)
y_grid <- seq(0, HEIGHT, step = DIAMETER)

draw_grid <- function() {
    for (x in x_grid) {
        for (y in y_grid) {
            # Uniform distribution for RGB color
            rgb_color = runif(3, 0, 255)
            fill(rgb_color)    # fill color
            stroke(rgb_color)  # border color
            
            # Square of normal distribution for the diameter
            d <- DIAMETER * rnorm(1, 0, 0.4)^2
            circle(x, y, d)
        }
    }
    NULL
}

setup <- function() {
    createCanvas(WIDTH, HEIGHT)
    draw_grid()
}

mouseClicked <- function() {
    clear()  # remember to clear the previous drawing
    draw_grid()
}
```

Click to update!

## Exercises

Try creating the following visualisation (which updates upon click):

```{r, echo = F, eval = TRUE}
sketch::insert_sketch("creative_coding_exercise.R",
                      asset_tags = default_tags(local = FALSE),
                      style = NULL,
                      class="border-none width-420 height-420")
```


## Credits / Reference

I learned the above visualisations from [Matt DesLauriers](https://github.com/mattdesl/workshop-generative-art). Visit the repository for other useful resources!


<link href="../css/docs-styles.css" rel="stylesheet"></link>

```{r, echo = F, eval = F}
# Do not use this block with docusaurus
source("highlight_Rmd.R")
highlight(query("pre code", 2), 17:24, include_source = T, 
          style = NULL, class = "highlight")
highlight(query("pre code", 3), c(13, 27, 31, 34:37), 
          style = NULL, class = "highlight")
```
